MLH1 Results Outline

### 2020-02-10 doc/Results.rmd
Cleaning up notes and better organization of this doc

- All the effects are unorder factors

@@

The fixed sex effect has the largest effect sizes. All are negative (for male).  (I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
The strain effects are only slightly significant for within mouse variance measures in the full dataset (includes low quality cells)

@@

For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
Most of the strain specific differences in within mouse variance go away in the Q12 dataset.

@@

# Figure 1

@@

# Proportion of chrm classes

TODO -- try running a chi square test on the proportions.

make a smaller proportion chrm block --

@@

This is also another way to score heterochiasmy (that gets around the sex chrms, i think)

@@

High musc males have half 1CO and 2CO bivalents.
For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).

Add p.values of tests for the how the proportions differ.

*cast female data came from another publication, add this ref (Kohler?)

The bottom lines and subspecies points don't exactly match up.

**add in cell images like the other draft figures.**

# Chrm Class proportions

The figure below shows the average percentages of chromosome classes per category.

@@

- Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.

# DMC1 Results Spermatocytes

@@

# Note on Heterochiasmy Definition

I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.

Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count). (I note if the and Y are paired, which they are at a high rate). The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).

To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.

@@

@@

The breakdown of curated bivalents by category

@@

@@

@@

As alex said, be careful about mixed models and interperting the random effect (radnom variance) -- assigning strain as random effect is slightly problamatic as i can give it a biological interpertation (it's due to strain -- genetics of each of the strains).

> Q1. same models for MLH1 based sex differences (and evolution)

> Q2 the same models for the male polymorphism and logistic regression for testing if bivalent traits can predcit if they are from high or low rec group

@@

- Simple application of the same models as MLH1 based reseults.

@@

```{r biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}

@@

```{r biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}

@@

@@

The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents. When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.

@@

# SC Lengths

We expect female SC lengths to be longer (refs).

In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.

For the 0 class chromosomes most all are around the same size of the 1CO distribution.

Add in the code for the plot under 2COs males and females

#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)

Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

The dependant variables I'll be testing in the mixed model framework are:

- long.biv

@@

Insert the results for the MMs using the dependant variable 1CO - use the Mixed model (lmer), for all of the meterics (pooled SC, 1CO 2CO ect)

Below's the code for Mixed model results -- try to organize them into a table or somethinf

```{r M1_many.metrics, echo=FALSE}
#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values

#all pooled SC lengths
lmer.SCLength_all_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#these 2 below are LR tests
drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
anova(lmer.SCLength_all_MM.M1) # also F values

exactRLRT(lmer.SCLength_all_MM.M1)
as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

#1CO
SCLength_1CO_MM.M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
drop1(SCLength_1CO_MM.M1, test="Chisq")
exactRLRT(SCLength_1CO_MM.M1)
as.data.frame(round(summary(SCLength_1CO_MM.M1)$coefficients, 6))

#2CO
SCLength_2CO_MM.M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
drop1(SCLength_2CO_MM.M1, test="Chisq")
exactRLRT(SCLength_2CO_MM.M1)
as.data.frame(round(summary(SCLength_2CO_MM.M1)$coefficients, 6))

#Long.Biv
SCLength_long.biv_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)

drop1(SCLength_long.biv_MM.M1, test="Chisq")
exactRLRT(SCLength_long.biv_MM.M1)
as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))

#### SC_HetC_M2

```{r M2_many.metrics, echo=FALSE, eval=FALSE}

#feed these 2 DFs into M2
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

#### SC_HetC_M3

```{r HetC_M3, echo=FALSE}

$$mouse \ average \ SC  \ length ~=~ strain + \varepsilon $$
M2 anova for glm, of each subset, does at least one of the strains differ from others?

M3 logistic regression, can a bivalent metric predict the high or low group?

$$ log( strain ~=~ \ mouse \ average \ SC  \ length ) $$

The IFD traits are:

- IFD_ABS

- IFD_PER

$$mouse \ average \ CO \ position \ trait  ~=~ rand(strain) + \varepsilon $$

These are the models for male polymorphism, (I am re-thinking the tests I think)

Use these DFs for the models:  **Male.poly.Mouse.Table_BivData_4MM**  and **Long.biv_mouse_table_male.poly**

** add chunks for glm()s of male polymorphism**

@@

### Summary

@@

> Is the random Effect of strain an effect on SC length?

The exactRLRT() test indicates that the random strain effect might be significant, p= xx

(merge this with the code above)
Prediction, High rec males have longer SC.

@@

Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))

#All chrm
log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_SC.length_mouse.av)

@@

#1CO
log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_SC.length_mouse.av_1CO)

@@

summary(log.reg_SC.length_mouse.av_2CO)

#3CO
log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_SC.length_mouse.av_3CO)

@@

summary(log.reg_SC.length_BIV)

@@

                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_SC.length_BIV_1CO)

#2CO
log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_SC.length_BIV_2CO)

@@

summary(log.reg_SC.length_BIV_3CO)
```

The mean SC logistic regresion model for mouse averages

@@

#### t.test for SC Length

@@

The table above should display the slightly unsual pattern, where the coefficients for the significant sex fixed effect are positive and neagtive in the raw and normalized values respectively. That is for the raw IFD values, females are signiciantly longer but the normalized IFD values, males are significantly longer.

Dive deeper into the sex specific pattern for each strain.  Below are code chunks which show the unusal sex specific results for IFD measures. The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD. The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.

  }
entr_DF

#uggh how does this not work ...

#more males have lower than 30% IFD?

#would Chi-sq be a good test to compare the ratios ect?

#the sum of these doesn't add

G.Bivalents.DF

```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
#real.long.bivData
#MELT_Curated_BivData

the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.

All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.

I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.

**What is the overlap for cell images in the MLH1 comparisons and these bivalent comparisons?**

should all mice be represented.

```{r male.Nrm1CO_strain, echo=FALSE}
#isolate all musc
Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

#make table
Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
                        n_obs = length(Obj.ID),
                         mean_F1 = mean(Foci1, na.rm=TRUE),
                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
                         )

#add rec group
mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )

## SC_male.poly_glm

remeber the 3 predictions for subsp;

glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
summary(glm1.SC.male_long.biv)

glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
summary(glm2.SC.male_long.biv)

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
summary(glm_SC.male_Musc_long.biv)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
summary(glm_SC.male_Mol_long.biv)

summary(log.SC.male_long.SC)

The glm for SC lengths -- sub-grouped by subsp -- indicates that all strain effects are significant for the average mouseisn't met.  SC length.
My prediction for Dom, isn't met.

(indecation that there is more variation for the SC lengths than for gwRR / CO counts)

-->
I think these comparisons should be done with the **Bivalent observations**
The logistic regression showed rec groups could be predicted by SC length - re-run them while seperating out by chrm class

The top t.tests indicate that when all the mice averages are pooled, there's a sig difference in SC lengths. But When the means are compared across chrm class, the mouse averages are no longer significant.

How well does SC length predict chromosome class?  Prediction, High rec strains will have more sig pvalue, given the lower overlap in SC lengths across chrm class. (Note this test can only be done with bivalent level observations, because 1CO or 2CO)

@@

- Is there a consistant pattern with the high rec strains?

@@

The t.tests return very significant pvalues for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).

@@

Examine the pattern in the Long bivalent dataset to get a feel for if the chromosome size effect skews the pattern. In the long bivalant data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5). The section on IFD only looks at the 2CO bivalents.

@@

There are alot fewer long bivalent data observations, but the same positive correlation with

@@

There are not enough observations for subsetting the data into strains or use the mouse averages.

Neither t.test are sig for both the ABS and PER when I test just the Musc strains.
The above t.tests are breaking the knitr

@@

put all of the code chuncks/analysis for caveats here

@@

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
--> 
Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plts excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correclty mapped)

@@

in code chunk above I ran the mouse averages for the longest bivalets. (680 bivalents, from 54 mice. 
10202 bivalents from 86 mice.

>>>>

- All the effects are un-ordered factors

@@

The fixed sex effect has the largest effect sizes. All are negative (for male).  (I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
The strain effects are only slightly significant for within mouse variance measures in the full data set (includes low quality cells)

@@

For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
Most of the strain specific differences in within mouse variance go away in the Q12 data set.

@@

<!--  come back to Fig1
# Figure 1

@@

# Proportion of chromosome classes

<!--
TODO -- try running a chi square test on the proportions.

make a smaller proportion chrm block --
Add p.values of tests for the how the proportions differ.
*cast female data came from another publication, add this ref (Kohler?)
The bottom lines and subspecies points don't exactly match up.
**add in cell images like the other draft figures.**
-->

@@

@@

For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).

The figure below shows the average percentages of chromosome classes per category gathered from BivData

@@

Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.

# DMC1 Results, Spermatocytes

@@

## Note on Heterochiasmy Definition

<!-- add outline for new XX adjust section 
adjust the colors for the plot below (standardize)
-->

I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison. In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape. Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count). (I note if the and Y are paired, which they are at a high rate). The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y). To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.

@@

(Put the XX adjustment section here)

@@

The breakdown of single bivalent observations by category

@@

Two Main questions for these analyses.

@@

These questions use different datasets, q1 sex matched strains

Q2 all male observations with including strains not in the q1 data set

@@

<!--
As alex said, be careful about mixed models and interpreting the random effect (random variance) -- assigning strain as random effect is slightly problematic as i can give it a biological interpretation (it's due to strain -- genetics of each of the strains).  
-->

Q1. Model 1, mixed model lmer()  
$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$

Q1. Model 2, glm()
$$mouse \ av.\  metric ~=~ sex * strain + \varepsilon $$

Q2 Model 1, mixed model lmer()

$$mouse \ av.\  metric ~=~ subsp *  + rand(strain) + \varepsilon $$

Q2 Model 2,  glm()

$$mouse \ av.\  metric ~=~ strain + \varepsilon $$

Q2 Model 3, logistic regression

$$Rec \ group  ~=~ mouse \ av.\  metric  $$

@@

- Simple application of the same models as MLH1 based results.

@@

```{r Q1.biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}

@@

```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}

@@

# Q1 SC Lengths

We expect female SC lengths to be longer (refs). In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number. For the 0 class chromosomes most all are around the same size of the 1CO distribution. Add in the code for the plot under 2COs males and females

#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

> M1. subsp * sex + (strain)

> M2 subsp \* sex \* strain  (don't rally use)

> M3 sex * strain

The dependent variables I'll be testing in the model framework are:

- Xlong.biv  *Don't do long bivData for testing sex effect, since this would have the XX

- Short.biv dataset *this would be a chrm size effect, without the XX

Insert the results for the MMs using the dependent variable 1CO - use the Mixed model (lmer), for all of the metrics (pooled SC, 1CO 2CO ect)

Below's the code for Mixed model results -- try to organize them into a table or something

```{r 1.SC.Length_M1, echo=FALSE}
#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values

#all pooled SC lengths
lmer.Q1_SCLength_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#these 2 below are LR tests
#drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
#anova(lmer.SCLength_all_MM.M1) # also F values
exactRLRT(lmer.Q1_SCLength_M1)
#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

#1CO  lmer.Q1_SCLength_M1
Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
#drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")
exactRLRT(Q1.lmer_SCLength_1CO_M1)
#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))

#2CO
Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
#drop1(SCLength_2CO_MM.M1, test="Chisq")
exactRLRT(Q1.lmer_SCLength_2CO_M1)
as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))

#Long.Biv
Q1.lmer_SCLength_long.biv_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)

#drop1(Q1.lmer_SCLength_long.biv_M1, test="Chisq")
#exactRLRT(Q1.lmer_SCLength_long.biv_M1)
#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
```

The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.

#### Q1. SC_HetC_M2

```{r Q1.SC.Length_M2, echo=FALSE}
#feed these 2 DFs into M2
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
#glm(metric ~ strain)  ---mean_SC

Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_SCLength_M2)

#Dom.all
Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.Dom)

#Dom.1CO
Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.1CO.Dom)

#Dom.2CO
Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.2CO.Dom)

#Musc.all
Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1glm_SCLength_M2.Musc)

#Musc.1CO
Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1glm_SCLength_M2.1CO.Musc)

#Musc.2CO
Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1glm_SCLength_M2.2CO.Musc)

#Mol.all
Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1glm_SCLength_M2.Mol)

#Mol.1CO
Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1glm_SCLength_M2.1CO.Mol)

#Mol.2CO
Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1glm_SCLength_M2.2CO.Mol)
```

```{r Q1.SC.M2plot, echo=FALSE}

#Q1 plots for SC Lengths across Musc strains

Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
#Musc.all
Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)

#mouse av
Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
Q1.SClength_Musc_plot

#BivData Level
Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")

Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))

Q1.SClength_Musc.BivData_plot

#BivData Level
Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")

Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))

Q1.SClength_Dom.BivData_plot

I think the strange thing about the Musc is that Skive females have shorter SC lengths.

The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively. That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.

Dive deeper into the sex specific pattern for each strain.  Below are code chunks which show the unusual sex specific results for IFD measures. The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD. The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.

  }
#entr_DF

#would Chi-sq be a good test to compare the ratios ect?
#the sum of these doesn't add 
#G.Bivalents.DF

```{r Q1.Nrm.F1pos, echo=FALSE}

#calq the means for normalized F1

```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}

#real.long.bivData
#MELT_Curated_BivData

#these isn't long.biv data
MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]

real.long.bivData
MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]

the mouse average foci1 pos is more significant in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.

All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.

@@

The mouse averages for the other position metrics will be highly influenced by proportions of the 1CO and 2CO bivalents. When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.

@@

@@

remember the 3 predictions for subsp;

```{r Q2.male.Nrm1CO_strain, echo=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

#add rec group
mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
```

glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
#summary(glm1.SC.male_long.biv)

glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
#summary(glm2.SC.male_long.biv)

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
#summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
#summary(glm_SC.male_Musc_long.biv)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
#summary(glm_SC.male_Mol_long.biv)

#summary(log.SC.male_long.SC)

Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.

The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).

The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class. Mainly that 1CO are shorter in the high rec strains then low rec strains. This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up. Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.  This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group. They have a lower probability of 2COs below a certain SC length threshold.

I think these comparisons should be done with the **Bivalent observations**
The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class

The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths. But When the means are compared across chrm class, the mouse averages are no longer significant.

How well does SC length predict chromosome class?  Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class. (Note this test can only be done with bivalent level observations, because 1CO or 2CO)

<!--
blocks of other male Q2 - SC lengths

@@

@@

@@

#male poly shouldn't be here
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))

#All chrm
log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av)

@@

#1CO
log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_1CO)

@@

#summary(log.reg_SC.length_mouse.av_2CO)

#3CO
log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_3CO)

@@

#summary(log.reg_SC.length_BIV)

@@

                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_1CO)

#2CO
log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_2CO)

@@

#summary(log.reg_SC.length_BIV_3CO)
```

The mean SC logistic regression model for mouse averages

@@

@@

@@

- Is there a consistent pattern with the high rec strains?

@@

The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).

@@

Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern. In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5). The section on IFD only looks at the 2CO bivalents.

@@

There are many fewer long bivalent data observations, but the same positive correlation with

@@

There are not enough observations for sub setting the data into strains or use the mouse averages.

Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
The above t.tests are breaking the knitr

@@

put all of the code chunks/analysis for caveats here

@@

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
--> 
Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plots excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correctly mapped)

@@

in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
10202 bivalents from 86 mice.
